Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-09 introduction.tex
Made description of confidence intervals more quantitative
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@992 ccdd05d5-56c1-461a-bf98-85f6876523b4
Polished caption for Figure 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@993 ccdd05d5-56c1-461a-bf98-85f6876523b4

We show that by examining the signal to noise ratio (\SNR{}) stream for
threshold crossings before the \GW{} signal leaves the detection band, it is
possible to trade some \SNR{} and sky localization
accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.  The
most likely estimates indicate that $\sim$10 sources will produce an \SNR\
above 8 in the detector 10 seconds or more prior to the merger during one
year of observation, though the uncertainty spans orders of magnitude.
The gray bands give the pessimistic to optimistic interval defined
in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
sources that will be detectable $t$ seconds before coalescence.  The heavy
solid line is the most likely yearly rate estimate $\dot N_{\mathrm{re}}$
during Advanced \LIGO.  The shaded region represents the confidence interval
$[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for detection and
that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a detector in
the `zero detuning, high power' configuration, $\sim$10~yr$^{-1}$ will be
detectable within 10 seconds before merger and $\sim$1~yr$^{-1}$ will be
detectable within 100 seconds of merger.} 
\end{center}
\end{figure}

>>>>

We show that by examining the signal to noise ratio (\SNR{}) stream for
threshold crossings before the \GW{} signal leaves the detection band, it is
possible to trade some \SNR{} and sky localization
accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
\LIGO\ assuming event rates predicted in~\cite{Abadie:2010p10836} and anticipated
detector sensitivity described in \cite{ALIGONoise}.  The
most likely estimates indicate that $\sim$10 sources will produce an \SNR\
above 8 in the detector 10 seconds or more prior to the merger during one
year of observation, though the uncertainty spans orders of magnitude.
The gray bands give the 5 to 95\% confidence interval as described
in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
sources that will be detectable $t$ seconds before coalescence.  The heavy
solid line is the most likely yearly rate estimate $\dot N_{\mathrm{re}}$
during Advanced \LIGO.  The shaded region represents the confidence interval
$[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  With an advanced detector in the `zero detuning,
high power' configuration \cite{ALIGONoise} and an \SNR\ threshold of 8,
we will observe a total of $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$, but
$\sim$10~yr$^{-1}$ will be detectable within 10~s of merger and
$\sim$1~yr$^{-1}$ within 100~s.} 
\end{center}
\end{figure}
